---
title: "STAC33-A10"
author: "Khusanbek Mukhammad Azim"
date: "08/04/2021"
output: html_document
---

```{r, message=FALSE}
library(tidyverse)
library(lubridate)
```


# Q.1

## (a)

Since the file is a .csv, we use read_csv.

```{r}
url <- "http://ritsokiguess.site/STAC33/sleeping.csv"
sleep <- read_csv(url)
sleep
```

We have 7 rows and 3 columns. bed.time and rise.time columns display the date and time in YMD HMS format. The sleep.time column displays the hours sleeping in decimal format.

## (b)

First the period that the student spends in bed in hours is computed.

```{r}
sleep %>% mutate(t = as.period(bed.time %--% rise.time)) %>% 
  mutate(time_in_bed = hour(t) + minute(t)/60) %>% 
  select(-t) -> sleep2
sleep2
```

The fractional time in hours that each student was in bed is given below.

```{r}
sleep2 %>% select(time_in_bed)
```

## (c)

We compute sleep efficiency be dividing the sleep time and time in bed and multiplying result by 100 to get a percentage.

```{r}
# compute sleep efficiency
sleep2 %>% 
  mutate(sleep_efficiency = (sleep.time/time_in_bed)*100) %>% 
  select(bed.time, sleep_efficiency)
```

The greatest sleep efficiency of 93.93% was achieved on the night of September 8, 2013.

## (d)

The time spend in bed formatted as H:M:S is given by determining the period from bed.time to rise.time.

```{r}
sleep %>% 
  mutate(time = as.period(bed.time %--% rise.time)) %>% 
  select(bed.time, time)
```

## (e)

```{r}
sleep %>% 
  mutate(bed_time = format(sleep$bed.time, format='%H:%M:%S')) %>% 
  tibble(Night = 1:7, time_in_bed=bed_time) %>% 
  ggplot(aes(x=Night, y=time_in_bed)) + geom_point() + scale_y_discrete(limits=rev) + ylab("Bed Time")
```

# Q.2

## (a)

```{r}
url <- "http://ritsokiguess.site/STAC32/heightfoot.csv"
ht <- read_csv(url)
ht
```

We have 33 rows and 2 columns.

## (b)

Height is the response variable and foot is the explanatory variable.

```{r}
# compute beta hat

x <- matrix(c(rep(1,33), ht$foot), nrow=33)
y <- ht$height

# calculate inverse
a <- solve(t(x)%*%x)

# calculate transpose of x time y
b <- t(x)%*%y

# beta hat
b_hat <- a%*%b
b_hat

```

The intercept is 34.336335 and the slope is 1.359062.

## (c)

```{r}
ht.lm <- lm(height~foot, data=ht)
summary(ht.lm)$coefficients
```

The correct intercept is 34.336335 and correct slope is 1.359062 which are identical to the calculated values.


